Update of the ozone dry deposition in the OsloCTM3 v1.0

### 2018-11-15 DryDep.tex
Update of some figures, correction of introduction text to Section 3. Tried to figure out about dry deposition onto deserts...

wherein aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989,ACP:Seinfeld2006}. For all gases, $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on land surface types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.). Originally, \citet{AE:Wesely1989} used fixed seasonal average dry deposition resistances for each land surface type. For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years. \citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances. An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{cm\,s^{-1}}$]{ACP:Helmig2007}. Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}. But progress has also been made on process oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}. The variety of differing formulations and choices of parameters leads to a wide spread in model intercomparisons \citep{ACP:Hardacre2015,AE:Derwent2018} and about 20\,\unit{\%} uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\

@@

\subsection{Dry deposition scheme for other gases and aerosols}
For \chem{NH_3}, \chem{SO_2}, and \chem{HNO_3}, we follow exactly the treatment as described by \citet{ACP:Simpson2012}. Since the dry deposition velocity for \chem{CO} is very small, we use a fixed value for $v_d^\chem{CO}$ of 0.03\,\unit{cm\,s^{-1}}.

@@

  a_1^\text{hydrophob.} = 0.025\,\unit{cm\,s^{-1}} \cdot \overline{u_*}

@@

  a_1^\text{hydrophil.} = 0.2\,\unit{cm\,s^{-1}} \cdot \overline{u_*},

@@

In this section, we present results from a manifold of OsloCTM3 model runs testing different parameters of the EMEP scheme. We focus on the changes in ozone, e.g., total dry deposition $\chem{O_3^\text{DD}}$, dry deposition velocities $v_\chem{O_3^\text{DD}}$, and surface concentration $[\chem{O_3}]_0$ respective atmospheric surface ozone burden in units of \unit{kg}, related to the stomatal conductance parameterization. We evaluate our results with respect to \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}). The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS version cy38r1nc4}. CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered by NASA's Global Fire Emissions Database, Version4 (GFEDv4). In the following, we will discuss the various model sensitivity studies.

@@

As presented in Section~\ref{sec:model_des}, the OsloCTM3 dry deposition scheme has undergone a major revision. Due to the previously generalized implementation of the Wesely scheme with respect to dry deposition resistances and surface types, it is not possible to fully disentangle the effects of every single change. Therefor, we have conducted 9 different model simulations to mainly assess the parameter space of the stomatal conductance within EMEP scheme. A reference simulation featuring the Wesely scheme has been done and will be refered to as \emph{Wesely type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models. For all model integrations, the meteorological reference year is 2005. This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years. After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} was supposed to be the baseline simulation for our sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}. As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively. During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions. Especially in Europe, this basically results in the subpression of canopy resistance in these regions. The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro}), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere (\emph{EMEP\_ppgssh}). Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3. Accidently, we used forcings for the year 2014 instead of our meteorological year 2005 and used this opportunnity to assess the impact on ozone regarding differing emissions. As mentioned previously, the simulations inititally ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$. A final sensityivity study was conducted changing to deep soil water content (\emph{EMEP\_SWVL4}).

@@

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    Simulation & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & 5\,\unit{^\circ C}-days & (\unit{cm\,s^{-1}}) & (year)\\
    \middlehline
    Wesely type       & x & \multicolumn{3}{c}{n/a}  & x &       & 1/2000  & 2014 \\
    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  %\belowtable{} % Table Footnotes
\end{table*}
%
In Fig.~\ref{fig:diff_maps}, we show the relative difference between \emph{EMEP\_ppgsssh} and the \emph{Wesely type} by means of surface ozone burden, dry deposition velocity and removal of ozone due to dry deposition. Surface ozone increases globally except for regions covered by tropical forest. Especially in desert regions in Africa and Asia, surface ozone increases by up to 100\,\unit{\%}. Consistently, dry deposition velocities decrease globally and up to 100\,\unit{\%} in desert regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only a small change.
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely type} in (a) total surface ozone; (b) average ozone dry deposition velosity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with recent multi-model comparison}
\label{subsec:model}
Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}. \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...

\begin{figure*}[t]
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep}\\
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydepvelo}
  \caption{Comparison of various model runs with the OsloCTM3 and the multi-model-mean results by \citet{ACP:Hardacre2015}. (a) total annual ozone drydeposition; (b) annual average ozone dry depositon velocity.}
  \label{fig:mmm_drydep}
\end{figure*}
%
\begin{figure*}[t]
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
  \caption{Seasonal cycle of total amount of ozone removed from the atmosphere through dry deposition seperated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). Comparison with results from \citet{ACP:Hardacre2015}.}
  \label{fig:mmm_drydep_hem}
\end{figure*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}
  \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocitys.}
  \label{fig:mmm_drydep_season}
\end{figure}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly_pft_nh}
  \caption{Average ozone dry deposition velocitys seperated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
  \label{fig:mmm_drydep_season_pft}
\end{figure}

>>>>

wherein aerodynamic $R_a$, quasi-laminar layer $R^i_b$, and canopy resistances $R^i_c$ are key components \citep{AE:Wesely1989,ACP:Seinfeld2006}. For all gases, $R_a$ is the same, while $R^i_b$ and $R^i_c$ vary from gas to gas and also depend on land surface types (e.g., ice/snow, water, urban, desert, agricultural land, deciduous forest, coniferous forest etc.). Originally, \citet{AE:Wesely1989} used fixed seasonal average dry deposition resistances for each land surface type. For all three types of resistances in this Wesely-type parameterization, more process-oriented formulations have been developed and validated over the years. \citet{ACP:Luhar2017} have validated ozone dry deposition to the ocean with respect to three different formulations of surface resistances. An update on the ozone surface resistance over snow and ice covered surfaces has been provided from combined model and observation studies \citep[][$v^\chem{O_3}_\text{ice/snow} = 1/10000\,\unit{m\,s^{-1}}$]{ACP:Helmig2007}. Canopy conductance is parameterized at the single-leaf-level (stomatal conductance) for various plant function types (PFT) as well as single species based on empirical studies \citep{PTRS:Jarvis1976, BallBerry1987, ACP:Simpson2012, ICP:MappingManual2017}. But progress has also been made on process oriented modeling of stomatal conductance \citep{AFM:Anderson2000,PP:Buckley2017}. The variety of differing formulations and choices of parameters leads to a wide spread in model intercomparisons \citep{ACP:Hardacre2015,AE:Derwent2018} and about 20\,\unit{\%} uncertainty on the resulting total dry deposition \citep{ACP:Monks2015}.\\

@@

\subsection{Dry deposition scheme for other gases and aerosols}
For \chem{NH_3}, \chem{SO_2}, and \chem{HNO_3}, we follow exactly the treatment as described by \citet{ACP:Simpson2012}. Since the dry deposition velocity for \chem{CO} is very small, we use a fixed value for $v_d^\chem{CO}$ of 0.03\,\unit{m\,s^{-1}}.

@@

  a_1^\text{hydrophob.} = 0.025\,\unit{m\,s^{-1}} \cdot \overline{u_*}

@@

  a_1^\text{hydrophil.} = 0.2\,\unit{m\,s^{-1}} \cdot \overline{u_*},

@@

In this section, we present results from a manifold of OsloCTM3 model integrations testing different parameters of the EMEP scheme. We focus on changes in ozone related to the stomatal conductance parameterization, e.g., total dry deposition $\chem{O_3^\text{DD}}$, dry deposition velocities $v_\chem{O_3^\text{DD}}$, and surface concentration $[\chem{O_3}]_0$ or surface ozone burden in units of \unit{kg}, respectivly. We evaluate our results with respect to the multi-model evaluation of ozone dry deposition by \citet{ACP:Hardacre2015} (Section~\ref{subsec:model}) and observations (Section~\ref{subsec:obs}). The OsloCTM3 is driven by meteorological input fields from \href{https://www.ecmwf.int/en/forecasts/documentation-and-support/evolution-ifs/cycle-38r1-summary-changes}{ECMWF -- OpenIFS version cy38r1nc4}. CEDS historical emission inventory is used for anthropogenic emissions, while biomass burning is covered in daily resolution by NASA's Global Fire Emissions Database, Version4 (GFEDv4). Biogenic emissions are taken from MEGAN output. In the following, we will present the various model sensitivity studies.

@@

As presented in Section~\ref{sec:model_des}, the OsloCTM3 dry deposition scheme has undergone a major revision. Due to the previously generalized implementation of the Wesely scheme with respect to dry deposition resistances and surface types, it is not possible to fully disentangle the effects of every single change. Therefor, the conducted sensitivity studies (9 in total) mainly assess the parameter space of the stomatal conductance within the EMEP scheme. A reference simulation featuring the Wesely scheme has been done and will be refered to as \emph{Wesely\_type}, indicating that other implementations of the original work by \citet{AE:Wesely1989} may exist in other models. For all model integrations, the meteorological reference year is 2005. This choice only affects the comparison with data and multi-model studies that either perform analysis on decadal averages or differing years. After finishing the work on the integration of the EMEP parameterization and initial testing, \emph{EMEP\_full} was supposed to be the baseline simulation for our sensitivity studies, \emph{EMEP\_offLight} and \emph{EMEP\_offPhen}. As indicated by the names, these scenarios are extreme cases, setting $f_\text{light}$ and $f_\text{phen}$ to 1, respectively. During analysis of the results, we found, that the SMOKE-BEIS parameterization of the greening season did not extend to the boreal and subarctic regions. Especially in Europe, this basically results in the subpression of canopy resistance in these regions. The following sensitivity study therefor comprises two different versions of pre-processed greening season (refer to Section~\ref{subsec:pre-pro} for details), mid and high latitudes in the northern hemisphere only (\emph{EMEP\_ppgs}) and for both, northern and southern hemisphere mid and high latitudes (\emph{EMEP\_ppgssh}). Building on the latter, we want to confirm the importance of the $v^\chem{O_3}_\text{ice/snow}$ update \citep{ACP:Helmig2007} in the Subarctic and Arctics, within the framework of the OsloCTM3 (\emph{EMEP\_ppgssh\_ice}). Accidently, we have used forcings for the year 2014 instead of 2005. Though, we take this as opportunnity to assess the impact on ozone regarding differing emissions (\emph{EMEP\_ppgs\_2005}). As mentioned previously, the simulations inititally ran with the uppermost soil water level (SWVL1) in the implementation of $f_\text{SW}$. A final sensitivity study was conducted changing to soil water level at depths deeper than 1\,\unit{m} (\emph{EMEP\_SWVL4}).

@@

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been decribed in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    Simulation & Wesely scheme & \multicolumn{3}{c}{EMEP scheme} & \multicolumn{2}{c}{Greening season} & $v^\chem{O_3}_\text{ice/snow}$ & Forcings\\
    & & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & 5\,\unit{^\circ C}-days & (\unit{m\,s^{-1}}) & (year)\\
    \middlehline
    Wesely\_type       & x & \multicolumn{3}{c}{n/a}  & x &       & 1/2000  & 2014 \\
    EMEP\_full        &   & incl. TR & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offLight    &   & incl. TR & 1 & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_offPhen     &   & 1        & x & SWLV1     & x &       & 1/2000  & 2014 \\
    EMEP\_SWVL4       &   & incl. TR & x & SWLV4     & x &       & 1/2000  & 2014 \\
    EMEP\_ppgs        &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2014 \\
    EMEP\_ppgs\_2005  &   & incl. TR & x & SWLV1     &   & NH    & 1/2000  & 2005 \\
    EMEP\_ppgssh      &   & excl. TR & x & SWLV1     &   & NH+SH & 1/2000  & 2014 \\
    EMEP\_ppgssh\_ice &   & excl. TR & x & SWLV1     &   & NH+SH & 1/10000 & 2014 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  %\belowtable{} % Table Footnotes
\end{table*}
%
In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{EMEP\_ppgsssh} and \emph{Wesely\_type} on global maps by means of surface ozone burden, dry deposition velocity and total ozone due to dry deposition. Surface ozone increases globally except for regions covered by tropical forest. Especially in desert regions in Africa and Asia, surface ozone increases by up to 100\,\unit{\%}. Consistently, dry deposition velocities decrease globally (up to 100\,\unit{\%}) in desert regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible. 
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-ozone_diff_maps}
  \caption{Relative difference between the sensitivity simulation \emph{EMEP\_ppgsssh} and the \emph{Wesely\_type} in (a) total surface ozone; (b) average ozone dry deposition velosity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Comparison with multi-model evaluation}
\label{subsec:model}
Compare to \citet{ACP:Hardacre2015} as well as \citet{ACP:Luhar2017}. \citep{AE:Derwent2018} Uncertainties in surface ozone sources and sinks larger than in formulations? -> Need to read this one...

\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-vmr_surfaceozone}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep}
  \caption{Comparison of the manifold of OsloCTM3 integrations with respect to (a) surface ozone concentrations; and in comparison with the multi-model evaluation by \citet{ACP:Hardacre2015}: (b) Annual average ozone dry depositon velocity; (c) Total annual ozone dry deposition. The different colors indicate sets of associated simulation.}
  \label{fig:mmm_drydep}
\end{figure}
%
\begin{figure*}[t]
  \includegraphics[width=12cm]{pictures/final-total_ozone_drydep_hem}
  \caption{Seasonal cycle of total amount of ozone removed from the atmosphere through dry deposition seperated into northern hemisphere (NH), tropics (TR), and southern hemisphere (SH). Comparison with results from \citet{ACP:Hardacre2015}.}
  \label{fig:mmm_drydep_hem}
\end{figure*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydep_monthly}\\
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly}
  \caption{Comparison of monthly zonal distributions to results from \citep{ACP:Hardacre2015}. (a) Total amount of ozone removed from the atmosphere; (b) average ozone dry deposition velocitys.}
  \label{fig:mmm_drydep_season}
\end{figure}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{pictures/final-total_ozone_drydepvelo_monthly_pft_nh}
  \caption{Average ozone dry deposition velocitys seperated by land use type and compared to results from \citep{ACP:Hardacre2015}.}
  \label{fig:mmm_drydep_season_pft}
\end{figure}
